
### ANLY 503 Assignment 6

### Geospatial Visualization


```{r include=FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(janitor)
library(tidyverse)
library(ggplot2)
library(sf)
library(spData, warn.conflicts = F, quietly = T)
library(maps, warn.conflicts = F, quietly = T)
library(viridis)
library(ggpubr)
library(cowplot)
library(stringr)
```




```{r include=FALSE}
library(reticulate)
py_install('pandas')
py_install('numpy')
py_install('matplotlib')
py_install('plotly')
```

```{python include=FALSE}
import pandas as pd
import numpy as np
import matplotlib.pyplot as plt
import plotly.express as px
import plotly.offline as py
```



```{python include=FALSE}
election_data = pd.read_csv('~/Desktop/503data/countypres_2000-2016.csv')
#election_data.head(5)
election_data = election_data[(election_data['year']==2008) | 
                             (election_data['year']==2012) | 
                             (election_data['year']==2016)]
election_data = election_data[(election_data['party']=='democrat') | 
                             (election_data['party']=='republican')]
election_data['won_proportion'] = election_data['candidatevotes']/election_data['totalvotes']
election_data = election_data[['year', 'state', 'state_po', 'FIPS', 'county', 'party', 'won_proportion']]
election_data = election_data.dropna(subset=['FIPS'])
election_data = election_data.astype({'FIPS': int})
election_data = election_data.astype({'FIPS': str})
election_data['FIPS'] = election_data['FIPS'].apply(lambda x: x.zfill(5))
election_data = election_data.reset_index(drop = True)
election_data.head(10)
#election_data.to_csv('new_election_data.csv')
```



```{python include=FALSE}
unemployment_data = pd.read_excel('~/Desktop/503data/Unemployment.xls', sheet_name = 'Unemployment Med HH Income', skiprows = 7)
unemployment_data.head(5)
unemployment_data = unemployment_data[['FIPStxt', 'Stabr', 'area_name', 
                                       'Unemployment_rate_2008', 'Unemployment_rate_2012', 'Unemployment_rate_2016']]
unemployment_data = pd.melt(unemployment_data, id_vars=["FIPStxt", 'Stabr', 'area_name'], var_name="year", value_name="unemployment_rate")
unemployment_data.year = unemployment_data.year.apply(lambda x: x[-4:])
unemployment_data = unemployment_data.astype({'FIPStxt': str})
unemployment_data['FIPStxt'] = unemployment_data['FIPStxt'].apply(lambda x: x.zfill(5))
unemployment_data = unemployment_data.astype({'year': str})
unemployment_data = unemployment_data.astype({'FIPStxt': str})
unemployment_data = unemployment_data.rename(columns={"FIPStxt": "FIPS"})
unemployment_data.head(5)
#unemployment_data.to_csv('new_unemployment.csv')
```




```{r include=FALSE}
new_election_data = read_csv('~/Desktop/503data/new_election_data.csv')
new_unemployment_data = read.csv('~/Desktop/503data/new_unemployment.csv')
new_unemployment_data$FIPS = as.character(new_unemployment_data$FIPS)
new_unemployment_data$FIPS = str_pad(new_unemployment_data$FIPS, 5, pad = "0")
```


```{r include=FALSE}
shape = st_read('~/Desktop/503data/tl_2019_us_county-1/tl_2019_us_county.shp')
tx_map = shape %>% filter(STATEFP =='48') %>% rename('FIPS'='GEOID')
ca_map = shape %>% filter(STATEFP =='06') %>% rename('FIPS'='GEOID')
pa_map = shape %>% filter(STATEFP =='42') %>% rename('FIPS'='GEOID')
```




#### Texas

```{r include=FALSE}

# TX Election Vote in 2008
election_tx_2008=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='TX') %>% filter(year=='2008') %>% filter(party=='democrat')
election_tx_2008 = tx_map %>% left_join(election_tx_2008[,c('FIPS',"won_proportion")],by='FIPS') 
election_tx_2008_map = ggplot(election_tx_2008) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in TX in 2008', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

# TX Unemployment in 2008
unemploy_tx_2008=new_unemployment_data %>% filter(Stabr=='TX') %>% filter(year=='2008')
unemploy_tx_2008=tx_map %>% left_join(unemploy_tx_2008, by='FIPS')
unemploy_tx_2008_map = ggplot(unemploy_tx_2008)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno", direction = -1)+labs(title='Unemployment Rate in Texas in 2008', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))



# TX Election Vote in 2012
election_tx_2012=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='TX') %>% filter(year=='2012') %>% filter(party=='democrat')
election_tx_2012 = tx_map %>% left_join(election_tx_2012[,c('FIPS',"won_proportion")],by='FIPS') 
election_tx_2012_map = ggplot(election_tx_2012) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in TX in 2012', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

# TX Unemployment in 2012
unemploy_tx_2012=new_unemployment_data %>% filter(Stabr=='TX') %>% filter(year=='2012')
unemploy_tx_2012=tx_map %>% left_join(unemploy_tx_2012, by='FIPS')
unemploy_tx_2012_map = ggplot(unemploy_tx_2012)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno", direction = -1)+labs(title='Unemployment Rate in Texas in 2012', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  
                                           aspect.ratio =1,
                                           axis.text=element_text(size=14),
                                           axis.title=element_text(size=16,face="bold"))

# TX Election Vote in 2016
election_tx_2016=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='TX') %>% filter(year=='2016') %>% filter(party=='democrat')
election_tx_2016 = tx_map %>% left_join(election_tx_2016[,c('FIPS',"won_proportion")],by='FIPS') 
election_tx_2016_map = ggplot(election_tx_2016) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in TX in 2016', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

# TX Unemployment in 2016
unemploy_tx_2016=new_unemployment_data %>% filter(Stabr=='TX') %>% filter(year=='2016')
unemploy_tx_2016=tx_map %>% left_join(unemploy_tx_2016, by='FIPS')
unemploy_tx_2016_map = ggplot(unemploy_tx_2016)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno", direction = -1)+labs(title='Unemployment Rate in TX in 2016', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

```


```{r fig1, fig.height = 30, fig.width = 25, fig.align = "center", echo=FALSE}
ggarrange(election_tx_2008_map, unemploy_tx_2008_map, election_tx_2012_map, unemploy_tx_2012_map, election_tx_2016_map, unemploy_tx_2016_map, ncol = 2, nrow=3, align = 'hv', widths = c(7,7,7,7,7,7))
```


Comments:

By looking at the Vote ratio for Democrat in TX over these three elections years, most counties in TX had low votes for Democrat, but the west and south counties had pretty high votes for Democrat. The unemployment rate were similar in these three years.

#### Californina

```{r include=FALSE}

# CA Election Vote in 2008
election_ca_2008 = new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CA') %>% filter(year=='2008') %>% filter(party=='democrat')
election_ca_2008 = ca_map %>% left_join(election_ca_2008[,c('FIPS',"won_proportion")],by='FIPS') 
election_ca_2008_map = ggplot(election_ca_2008) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in CA in 2008', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"), legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# CA Unemployment in 2008
unemploy_ca_2008=new_unemployment_data %>% filter(Stabr=='CA') %>% filter(year=='2008')
unemploy_ca_2008=ca_map %>% left_join(unemploy_ca_2008, by='FIPS')
unemploy_ca_2008_map = ggplot(unemploy_ca_2008)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in CA in 2008', fill='Unemployment Rate(%)') + 
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"), aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# CA Election Vote in 2012
election_ca_2012 = new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CA') %>% filter(year=='2012') %>% filter(party=='democrat')
election_ca_2012 = ca_map %>% left_join(election_ca_2012[,c('FIPS',"won_proportion")],by='FIPS') 
election_ca_2012_map = ggplot(election_ca_2012) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in CA in 2012', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"), legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# CA Unemployment in 2012
unemploy_ca_2012=new_unemployment_data %>% filter(Stabr=='CA') %>% filter(year=='2012')
unemploy_ca_2012=ca_map %>% left_join(unemploy_ca_2012, by='FIPS')
unemploy_ca_2012_map = ggplot(unemploy_ca_2012)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in CA in 2012', fill='Unemployment Rate(%)') + 
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"), aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# CA Election Vote in 2016
election_ca_2016 = new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='CA') %>% filter(year=='2016') %>% filter(party=='democrat')
election_ca_2016 = ca_map %>% left_join(election_ca_2016[,c('FIPS',"won_proportion")],by='FIPS') 
election_ca_2016_map = ggplot(election_ca_2016) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in CA in 2016', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"), legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# CA Unemployment in 2016
unemploy_ca_2016=new_unemployment_data %>% filter(Stabr=='CA') %>% filter(year=='2016')
unemploy_ca_2016=ca_map %>% left_join(unemploy_ca_2016, by='FIPS')
unemploy_ca_2016_map = ggplot(unemploy_ca_2016)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in CA in 2016', fill='Unemployment Rate(%)') + 
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), 
                                           panel.spacing.x=unit(5, "lines"), aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

```


```{r fig2, fig.height = 30, fig.width = 25, fig.align = "center", echo=FALSE}
ggarrange(election_ca_2008_map, unemploy_ca_2008_map, election_ca_2012_map, unemploy_ca_2012_map, election_ca_2016_map, unemploy_ca_2016_map, ncol = 2, nrow=3, align = 'hv', widths = c(7,7,7,7,7,7))
```


Comments:

By looking at the Vote ratio for Democrat in CA over these three elections years, most counties in CA had high votes for Democrat, but the east and north counties had low votes for Democrat. The unemployment rate was higher in 2012 thann 2008 and 2016.


#### Pennsylvania

```{r include=FALSE}

# PA Election Vote in 2008
election_pa_2008=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='PA') %>% filter(year=='2008') %>% filter(party=='democrat')
election_pa_2008 = pa_map %>% left_join(election_pa_2008[,c('FIPS',"won_proportion")],by='FIPS') 
election_pa_2008_map = ggplot(election_pa_2008) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in PA in 2008', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title=element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# PA Unemployment in 2008
unemploy_pa_2008=new_unemployment_data %>% filter(Stabr=='PA') %>% filter(year=='2008')
unemploy_pa_2008=pa_map %>% left_join(unemploy_pa_2008, by='FIPS')
unemploy_pa_2008_map = ggplot(unemploy_pa_2008)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in PA in 2008', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))



# PA Election Vote in 2012
election_pa_2012=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='PA') %>% filter(year=='2012') %>% filter(party=='democrat')
election_pa_2012 = pa_map %>% left_join(election_pa_2012[,c('FIPS',"won_proportion")],by='FIPS') 
election_pa_2012_map = ggplot(election_pa_2012) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in PA in 2012', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title=element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# PA Unemployment in 2012
unemploy_pa_2012=new_unemployment_data %>% filter(Stabr=='PA') %>% filter(year=='2012')
unemploy_pa_2012=pa_map %>% left_join(unemploy_pa_2012, by='FIPS')
unemploy_pa_2012_map = ggplot(unemploy_pa_2012)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in PA in 2012', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

# PA Election Vote in 2016
election_pa_2016=new_election_data %>% select(c("year", "FIPS","state_po","party", "won_proportion"))%>% filter(state_po=='PA') %>% filter(year=='2016') %>% filter(party=='democrat')
election_pa_2016 = pa_map %>% left_join(election_pa_2016[,c('FIPS',"won_proportion")],by='FIPS') 
election_pa_2016_map = ggplot(election_pa_2016) +geom_sf(aes(geometry=geometry,fill=won_proportion))+
  scale_fill_viridis(option = "plasma", direction = -1)+labs(title='Democrat Vote in PA in 2016', fill='Vote Proportion')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title=element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))


# PA Unemployment in 2016
unemploy_pa_2016=new_unemployment_data %>% filter(Stabr=='PA') %>% filter(year=='2016')
unemploy_pa_2016=pa_map %>% left_join(unemploy_pa_2016, by='FIPS')
unemploy_pa_2016_map = ggplot(unemploy_pa_2016)+geom_sf(aes(geometry=geometry,fill = unemployment_rate))+
  scale_fill_viridis_c(option = "inferno",direction = -1)+labs(title='Unemployment Rate in PA in 2016', fill='Unemployment Rate(%)')+
  xlab('Longitude')+ylab('Latitude')+theme(plot.title = element_text(color = "black", size = 24, vjust = 1,hjust = 0.5, face="bold"),legend.title = element_text(color = "black", size = 18),
                                           legend.text = element_text(color = "black", size = 18), panel.spacing.x=unit(5, "lines"),  aspect.ratio = 1,
                                           axis.text=element_text(size=14),axis.title=element_text(size=16,face="bold"))

```


```{r fig3, fig.height = 30, fig.width = 25, fig.align = "center", echo=FALSE}
ggarrange(election_pa_2008_map, unemploy_pa_2008_map, election_pa_2012_map, unemploy_pa_2012_map, election_pa_2016_map, unemploy_pa_2016_map, ncol = 2, nrow=3, align = 'hv', widths = c(7,7,7,7,7,7))
```



Comments:

By looking at the Vote ratio for Democrat in PA over these three elections years, more counties in PA had high votes for Democrat in 2008 than 2012,The unemployment rate was higher in 2016 thann 2008 and 2012.


